---
title: "p8105_hw3_sl4471"
author: Shuwei Liu sl4471
date: 10.13.2018
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 1.5,
  out.width = "90%")
library(tidyverse)
library(ggridges)
library(patchwork)
library(hexbin)
```

# Problem1

## 1.1 Import and Clean the data

```{r import data}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data("brfss_smart2010")
overall_health_df = 
  janitor::clean_names(brfss_smart2010) %>%
  rename(state = "locationabbr") %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = factor(response, levels = str_c(
    c("Excellent", "Very good", "Good", "Fair", "Poor"))))
```

## 1.2 In 2002, which states were observed at 7 locations?

```{r 2002_data}
data_2002 = 
  filter(overall_health_df, year == "2002") %>% 
  group_by(state) %>% 
  summarize(n_obs = n(),
            n_loc = n_distinct(locationdesc)) %>% 
  select(state, n_loc) %>% 
  filter(n_loc == "7")
```

In the year 2002, state Connecticut, Florida and North Carolina were observed at 7 locations.

## 1.3 Spaghetti plot

```{r spaghetti fig.width = 7, fig.asp = 0.6, out.width = "90%"}
number_locations_df = 
  overall_health_df %>% 
  group_by(state, year) %>% 
  summarise(n_obs = n(), n_loc = n_distinct(locationdesc))
ggplot(number_locations_df, aes(x = year, y = n_loc, color = state)) +
  geom_line() +
  labs(
    title = "Number of locations in each state from 2002 to 2010",
    x = "number of locations",
    y = "year") +
  theme_bw() +
  theme(legend.position = "bottom")
```

I chose "geom_line" for the "spaghetti plot" since we need to show the changes of the number of locations in different states over years. The line plot could not only display the trend for changes of each states but also compare the number of locations between the states. From the plot, we can tell that most of the states have a stable change and some of them did not have a different number of locations over years. Florida has a disctint change of the number.

## 1.4 Table for the proportion of "Excellent" in NY
```{r excellent in NY}
NY_data = 
overall_health_df %>% 
  group_by(year) %>%
  filter(state == "NY", response == "Excellent", year %in% c(2002, 2006, 2010)) %>% 
  summarise(excellent_mean = mean(data_value, na.rm =  TRUE), 
            excellent_sd = sd(data_value, na.rm =  TRUE)) %>% 
  knitr::kable(digits = 1)
NY_data
```

From above, the mean of the proportion of "Excellent" response has decreased from 24% to 22.7%. However, the standard deviation of it also decreased which means that the range of the data for 2010 is smaller then the range of that for 2002. 

## 1.5 The distribution of state-level averages over time

```{r average_distribution fig.width = 6, fig.asp = 0.6, out.width = "90%"}
avearge_proportion_df = 
  overall_health_df %>% 
  group_by(state, year, response) %>%
  summarise(mean = mean(data_value, na.rm = TRUE))
ggplot(avearge_proportion_df, aes(x = year, y = mean, color = state)) +
  geom_line() +
  facet_grid(~response) +
  labs(
    title = "State-level average of response VS year",
    x = "Year",
    y = "Average proportion",
    caption = "Data from p8105.datasets") +
  theme_bw() +
  theme(legend.position = "bottom")
```
 
Each state has fairly same average proportion for each response. The "Very good" Response has the highest proportion while the "Poor" response has the least. 

## Problem2

## 1.1 Description of dataset


```{r import_data}
data("instacart")
str(instacart)
dim(instacart)
```

We can use "str()" and "dim" to look through the dataset. This a "tbl_df" data with 1384617 rows and 15 columns. It includes "order_id", "product_id", "add_to_cart_order", "reordered", "user_id", "eval_set", "order_number", "order_dow", "order_hour_of_day", "days_since_prior_order", "product_name", "aisle_id", "department_id", "aisle" and "department" variables. Basically, it tells us "who" order "what" kind of product under "which" aisle and department if it is reordered or not. To be specific, just take the first row as an example, the order id "1" shows that  customer "112108" reordered "Bulgarian Yogurt" whose id is "49302" under the yogurt aisle whose id is "120" from the "dairy eggs" department whose id is "16" on "Thursday" as this customers first order this time. 
```{r}
count(distinct(instacart, aisle))
instacart %>% 
  janitor::clean_names() %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))
```


